---
title: "Homework5"
author: Merilyn Thomas, Bryan Sandor, Tyler White
format: html
---

# Homework 5

## Task 1: Conceptual Questions

1.  The purpose of `lapply` is to apply a function to each element of the supplied list. The equivalent `purr` function is `map`
2.  `lapply(my_list, FUN = cor, method = "kendall")`
3. `purr` functions work within the tidyverse (they are pipeable) as well as having consistent arguments across different types that we are working with
4. A side-effect function is one that performs a certain task outside of the code (plot, print, generate csv, etc) and leaves the original data as-is
5. sd is defined locally within the function, so if that function is being called, R looks within that function's namespace first. If R does not find a variable called sd() with a function, it will use whatever is defined globally.


## Task 2

### Part 1

```{r}
getRMSE = function(y, yHat, ...){
  
  devs = (y - yHat)^2
  RMSE = sqrt(mean(devs, ...))
  
  return(RMSE)
  
}
```

### Part 2

Run example code
```{r}
set.seed(10)
n <- 100
x <- runif(n)
resp <- 3 + 10 * x + rnorm(n)
pred <- predict(lm(resp ~ x), data.frame(x))
```

Test
```{r}
getRMSE(y = resp, yHat = pred)
```

Test without specifying NA behavior
```{r}
resp[c(23,15)] = NA_real_

## Without NA Behavior
getRMSE(y = resp, yHat = pred)
```

Test With NA Behavior
```{r}
resp[c(23,15)] = NA_real_

## Without NA Behavior
getRMSE(y = resp, yHat = pred, na.rm = TRUE)
```
### Part 3
```{r}
getMAE = function(y, yHat, ...){
  absDev = abs(y - yHat)
  
  MAE = (mean(absDev, ...))
  return(MAE)
}
```

### Part 4

```{r}
set.seed(10)
n <- 100
x <- runif(n)
resp <- 3 + 10 * x + rnorm(n)
pred <- predict(lm(resp ~ x), data.frame(x))
```

Test without removing variables
```{r}
getMAE(y = resp, yHat = pred)
```
Remove some responses:
```{r}
resp[c(7,11)] = NA_real_
```

Test without NA behavior
```{r}
getMAE(y = resp, yHat = pred)
```

Test with NA behavior
```{r}
getMAE(y = resp, yHat = pred, na.rm = TRUE)
```

### Part 5
```{r}

getErrors = function(y, yHat, output = "all", ...){
  

    if (is.atomic(y) & is.numeric(y) & is.atomic(y)){
      invisible(y)
    }
    else{
      stop("Bad input")
      
    if (is.atomic(yHat) & is.numeric(yHat) & is.atomic(yHat)){
      invisible(y)
    }
    else{
      stop("Bad input")}
    
    
    
    
  }
  
  if (output == "RMSE"){
    return(c("RMSE" = getRMSE(y, yHat, ...)))
  }
  else if (output == "MAE"){
    return(c("MAE" = getMAE(y, yHat, ...)))
  }
  else if(output == "all"){
    RMSE = c("RMSE" = getRMSE(y, yHat, ...))
    MAE = c("MAE" = getMAE(y, yHat, ...))
    return(list(RMSE = RMSE, MAE = MAE))
  }
  else{
    stop("Bad Input")
  }
}

```


### Part 6

Generate Test Data
```{r}
set.seed(10)
n <- 100
x <- runif(n)
resp <- 3 + 10 * x + rnorm(n)
pred <- predict(lm(resp ~ x), data.frame(x))
```

Test function

Both MAE and RMSE
```{r}
getErrors(y = resp, yHat = pred)
```
MAE
```{r}
getErrors(y = resp, yHat = pred, output = "MAE")
```

RMSE
```{r}
getErrors(y = resp, yHat = pred, output = "RMSE")
```

Remove Data & Test
```{r}
resp[c(7,11)] = NA_real_
```

Test without NA behavior
```{r}
getErrors(y = resp, yHat = pred, output = "RMSE")
```
Test with NA Behavior
```{r}
getErrors(y = resp, yHat = pred, output = "RMSE", na.rm = TRUE)
```

Test with poorly specified data:
```{r}
getErrors(y = as.character(resp), yHat = pred, output = "RMSE", na.rm = TRUE)

```
```{r}
getErrors(y = data.frame(resp), yHat = pred, output = "RMSE", na.rm = TRUE)
```

