---
title: "Homework5"
author: Bryan Sandor, Merilyn Thomas, Tyler White
format: html
editor:
  markdown:
    wrap: 80
editor_options: 
  chunk_output_type: console
---

# Preamble
This homework explores advanced function writing in `R`.

# Task 1: Conceptual Questions
1. What is the purpose of the `lapply()` function? What is the equivalent `purrr` function?

2. Suppose we have a list called `my_list`. Each element of the list is a numeric data frame (all columns are numeric). We want to use `lapply()` to run the code `cor(numeric_matrix, method = "kendall")` on each element of the list. Write code to do this below! (I'm really trying to ask you how you specify `method = "kendall"` when calling `lapply()`)

3. What are two advantages of using `purrr` functions instead of the `BaseR` apply family?

4. What is a side-effect function?

5. Why can you name a variable `sd` in a function and not cause any issues with the `sd` function?

# Task 2 - Writing R Functions
## Subtask 1
For a set of responses $y_1, y_2, \ldots, y_n$ (variables of interest which we want to predict) and a set of corresponding predictions for those observations $\hat{y}_1, \hat{y}_2, \ldots, \hat{y}_n$, the Root Mean Square Error (RMSE) is $$RMSE = \sqrt{\frac{1}{n} \sum_{i = 1}^n (y_i - \hat{y}_i)^2}$$ Write a basic function that takes in a _vector_ of responses and a _vector_ of predictions and outputs the RMSE.

```{r subtask_1}
getRMSE <- function(vec_resp, vec_pred, ...){
  MSE <- mean((vec_resp - vec_pred)^2, ...)
  
  return(sqrt(MSE))
}
```

## Subtask 2
We run the following code to create some (controlled) response and prediction values.

```{r subtask_2_a}
#| code-overflow: wrap
#| 
set.seed(10)
n <- 100
x <- runif(n)
resp <- 3 + 10 * x + rnorm(n)
pred <- predict(lm(resp ~ x), data.frame(x)) # Note the tilde character ~ within the `lm` function
```

Now we test the RMSE function on the data.

```{r subtask_2_b}
getRMSE(resp, pred)
```

Now we test the RMSE function with two `NA_real_` values in the response data.

```{r}
resp[1] <- NA_real_
resp[2] <- NA_real_

getRMSE(resp, pred)

getRMSE(resp, pred, na.rm = TRUE)
```

## Subtask 3
Another common metric for evaluating predictions is the Mean Absolute Error (MAE) $$MAE = \frac{1}{n} \sum_{i = 1}^n |y_i - \hat{y}_i|$$ So we now write a function `getMAE()` providing the MAE for supplied data.

```{r subtask_3}
getMAE <- function(vec_resp, vec_pred, ...){
  return(mean(abs(vec_resp - vec_pred), ...))
}
```

## Subtask 4
We run the same randomizer as before

```{r subtask_4_a}
set.seed(10)
n <- 100
x <- runif(n)
resp <- 3 + 10 * x + rnorm(n)
pred <- predict(lm(resp ~ x), data.frame(x))
```

And test our new function on it

```{r subtask_4_b}
getMAE(resp, pred)
```

And finally check it with the optional argument

```{r subtask_4_c}
resp[1] <- NA_real_
resp[2] <- NA_real_

getMAE(resp, pred)

getMAE(resp, pred, na.rm = TRUE)
```

## Subtask 5
Now we construct a __wrapper__ function, utilizing the previous two functions.

```{r subtask_5}
wrapper <- function(x, y, RMSE = TRUE, MAE = TRUE, ...){
  if (is.numeric(x) & is.vector(x) & is.atomic(x) & 
      is.numeric(y) & is.vector(y) & is.atomic(y)){
    if (RMSE){
      cat(paste0("RMSE: ", getRMSE(x, y, ...)))
      if (MAE){
        return(cat(paste0("\n", "MAE: ", getMAE(x, y, ...))))
      }
    }
    if (MAE){
      cat(paste0("MAE: ", getMAE(x, y, ...)))
    }
  } else {
    print("Please supply two numeric (atomic) vectors!")
  }
}
```


## Subtask 6
Finally, we generate the pseudo-random data again

```{r subtask_6_a}
set.seed(10)
n <- 100
x <- runif(n)
resp <- 3 + 10 * x + rnorm(n)
pred <- predict(lm(resp ~ x), data.frame(x))
```

- We now call the new function thrice, specifying each metric individually and then together

```{r subtask_6_b}
wrapper(resp, pred, MAE = FALSE)
wrapper(resp, pred, RMSE = FALSE)
wrapper(resp, pred)
```

- Then we replace two of the values in the responses with `NA` and test the function

```{r subtask_6_c}
resp[1] <- NA_real_
resp[2] <- NA_real_

wrapper(resp, pred)
wrapper(resp, pred, na.rm = TRUE)
```

- Finally, we feed the wrapper function data that is not the correct format

```{r subtask_6_d}
a <- list(1, 2, "3")
b <- matrix(1:9, nrow = 3, ncol = 3, byrow = TRUE)
d <- c("1", "2", "3")

wrapper(a, b)
wrapper(b, d)
```

